\name: A non-parametric, hierarchical model for a precision spectrograph

### 2020-04-25 paper/paper.tex
Abstract Typo
Paper Re-Framing

\begin{abstract}
\name is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.  There have been many recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.  In such stabilized spectrographs, the full optical system and detectors will only vary within a tiny range of configurations.  Additionally, within a given range of stability, every calibration image ever taken contains information relevant to every science exposure ever taken.  The development of laser-frequency combs or etalons allows the spectral range of an instrument to be covered with dense calibration points.  Here, we describe a calibration method---\name---that is designed for this new era of stabilized instruments with high-line-density calibrators.  We find that a high density of calibration lines allows for the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.  We can also denoise the available calibration data by using it to determine not just the wavelength solution, but also the possible space of wavelength solutions, which is small for a stabilized instrument.  We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.  We find that \name\ returns more accurate wavelength predictions than classic polynomial models.
\end{abstract}

@@

\section{Introduction}

Extreme precision radial-velocity (\eprv) programs have been outrageously successful in finding and characterizing extra-solar planets (CITE THINGS).  
These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.  
Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).  
This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrograph is well sampled).  
Some hardware--software systems seem to be reaching these levels (CITE THINGS) .

@@

The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).  We expand on the work of Milakovi\'{c}2020, who proposed constructing a wavelength solution as multiple, segmented polynomials across an order by proposing a wavelength solution that is completely non-parametric.  \lz{(Should work in their paper more and praise it more?)}
Going non-parametric will improve calibration accuracy by not forcing the chose of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).

The second idea flows from the observation that contemporary
\eprv\ instruments are incredibly stable.
Temperature-controlled, fiber-fed pectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.
This renders it inappropriate to fit each calibration exposure or calibrate each science exposure independently.
Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.  Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed for each exposure.

This structure is \emph{hierarchical}: the calibration data are used not just to determine the wavelength solution, but also to determine the possible space of wavelength solutions.
In the statistics literature this concept is often described as \emph{de-noising}.  
Every calibration exposure contains information about every other exposure.  Thus every exposure can be improved (de-noised) with information coming from every other exposure.

The method we propose here---\name---embodies these ideas.
It is a non-parametric, hierarchical, data-driven model for the wavelength solution.
By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the observed variations in the spectrograph.
\name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, etalons, or good arc lamps.  The scope of the method is to take given calibration line positions, and de-noise/interpolate them into a full wavelength solution.
We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.  The motivating ideas behind this project are true for almost every astronomical spectrograph.

\section{Method} \label{sec:method}
We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized, spectrograph hardware.

We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ represents the parameters describing a given exposure.

A stabilized spectrograph should experience only low-degree variability, meaning the calibration of any image can be informed by the calibration of every other image.  To implement such a hierarchical model, we can use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.

If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.

@@

\subsection{Data} \label{sec:data}
We present results using data from \expres\, the EXtreme PRecison Spectrograph.  \expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).  \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g. \lz{Wilken2012, Molaro2013, Probst2014}).  ThAr exposures are taken at the beginning and end of each night.  LFC exposures are interspersed between science exposures every ~15-30 minutes.

The results and discussion presented here are based off of 1227 LFC exposures taken over two months months on 29 unique days.  LFC lines cover 40 echelle orders, which contain 19203 lines.  Though we primarily work with LFC data, we at times discuss applications to arc lamps.  These results use 78 ThAr exposures that were taken over the same range of time.  ThAr lines cover all 86 extracted orders of \expres\ and includes 5295 lines.

For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.  Gaussian profiles are fit to each line in the 1D, extracted data.  The mean of each fitted Gaussian is taken to be the center of the line (see \lz{Petersburg2020} for more details).  While precisely defining and determining line positions is essential to precise wavelength calibration, we consider the task out of scope for this paper.  \name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.

\expres\ exposures are separated into different ``instrumental epochs," which correspond to atypical, hardware changes that affected the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.  Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.

\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
\lz{Concerns: 1) how to best phrase the problem using variables?, 2) how in-depth is too in-depth?}

Let $\{exp_n\}$ be the set of calibration exposures within an instrumental epoch.  These exposures will be used in constructing the fiducial calibration of the spectrograph, $g_0(x,y,m)$, and the basis functions defining the spectrograph's calibration space, $g_k(x,y,m)$ (see equation 2).  We first identify the full list of lines we can expect to find in a calibration image from this epoch.  A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength assigned to that line, $\lambda$.  Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.

Each line, $l_{m,\lambda}$, has a fitted line center, $x_{n,l}$, in pixel space, for each calibration exposure $exp_n$.  If this line is missing from an exposure (e.g. the fit failed due to noise, the line isn't in its usual order, etc.), we set the line center $x_{n,l} = NaN$ for that line center in that exposure.  This allows us to construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.

Lines that are missing from greater than some percent of exposures are completely cut from the analysis.  Exposures that are missing greater than some percent of lines are also cut.  We use a threshold percentage of 50\% for both lines and exposures.  This largely affects the first lines on either end of an order and exposures with spuriously low signal.  Typically, about 0.2\% of lines and cut for LFC based calibration and 0.25\% of files are cut.

Missing line center measurements are replaced with denoised estimates using iterative PCA.  For each missing line center measurement, we initialize its value to the mean of the nearest $h$ exposures in time, where $h$ is typically $9$ exposures.  We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, which is set to be the mean line center value for each identified line over all exposures.  We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.  The resulting principal components are used as the basis functions defining the spectrograph's calibration space, $g_k(x,y,m)$ while the amplitude of each principal component is $a_{nk}$.

Using those values with equation 2 and some small integer (e.g. 6) for $K$, we can reconstruct a denoised matrix of line centers for each exposure.  Missing line center measurements are replaced with values from this reconstruction.  We iterate until the values of the missing line centers change by less than 0.01\% of a pixel.  This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."   We have found this typically takes less than 30 iterations.

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $\{x_{n,l}\} $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
\While{change in missing or outlier line centers $>$ 0.01\%}{
	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ for $K=6$\;
	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ was initially $NaN$
	}
\caption{Hierarchical De-Noising}
\end{algorithm}

Following the de-noising process, we check for deviant line centers.  This is aimed at catching lines that have been mis-identified or poorly fit, resulting in erroneous line centers.  Outliers are identified via a simple sigma cut for each line.  Any line center that is reported at more than $3\sigma$ away from the mean line center value is relabeled a ``bad line.''  We then repeat the process of hierarchical de-noising to also replace the line centers of these bad lines. (\lz{Should probably add something about why the sigma cut works?})

\subsection{Non-Parametric Calibration} \label{sec:nonparam}
The process of hierarchical de-noising gives us the amplitude for each principal component, $a_{nk}$, which captures the variability of the instrument.  We can interpolate these amplitudes $a_{nk}$ with respect to housekeeping data to determine line positions, $\lambda(x,y,m,n)$, for any state of the spectrograph.  In this paper, we show results where these amplitudes $a_{nk}$ are interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.

Once we have the line center for every line $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate  the known wavelengths over line centers to any position in x in a given order.  For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: $g_0(x,y,m)$; $a_{n,k}$; $g_k(x,y,m)$; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
Interpolate $a_{n,k}$ with respect to time to determine $a_{n,k}'$ for time $t'$\;
Calculate pixel line positions for time $t'$ by evaluating\;
$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ for $K=6$\;
\For{each unique $m$}{
	interpolate ${l_n,m}$ onto pixels that we want wavelengths for\;
	}
\caption{Non-Parametric Wavelength Solution}
\end{algorithm}

\section{Tests}
\subsection{Self Tests} \label{sec:test-self}
For these so-called ``self tests", we use \name\ to  predict wavelengths for the measured line centers of a calibration exposure that was used to construct the basis.  The linear interpolation in time should return identical PCA coefficients for this time, as the calibration data was used in the PCA.

The predicted wavelengths for these measured line centers should match the assigned wavelengths for these calibration lines within the range of error of the measured line center.  Per the EXPRES pipeline, the LFC line centers used in this analysis are measured by fitting each LFC line to a Gaussian \lz{Peteresburg2020}.  The mean of the fitted Gaussian is set to be the line center.  The error in this fitted value is on average 2.1 $\mps$.  Additional error will arise from interpolation error of wavelengths to other pixel positions.

\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
A more full-scale test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis vectors and weights with PCA.  We can then use the resulting basis to predict wavelengths for lines in the calibrations images that were left out.  The used calibration images represent the ``training set," while the left out exposures make up the ``validation set."

For these tests, we used 90\% of available exposures to train the PCA, and left the remaining 10\% for validation.  Exposures were sorted into either set at random.  Contrary to the self tests, these results may also accrue errors from interpolating the principal component amplitudes with respect to time, i.e. we are now also testing how well we can re-construct line centers for all times.   As before, differences in wavelength will also be drawn from how well line centers are measured in an exposure and any error in interpolating to those measured values.  This second source of error was independently quantified in the second test described in the previous section.

\lz{I had some discussion of cross tests (i.e. using LFC to predict ThAr and ThAr to predict LFC, but I removed it since we decided to leave ThAr mostly out of this due to its data generally being worse}

\section{Results} \label{sec:results}

\begin{figure}[h]
\centering
\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
\caption{Residuals of different wavelength calibration tests.  Shown as a black outline is the distribution of residuals from the self test described in Section \ref{sec:test-self}, which incorporates errors in the fitted line centers and interpolating with respect to pixel.  Over-plotted in blue are the residuals from the training and validation test described in Section \ref{sec:test-trainNvalid}, which builds on the self tests by additionally incorporating possible errors from interpolating line positions through time.\
\lz{Considering adding a histogram of line center errors to show it is on par?  i.e. the limitation is the line fit.  What other information is needed?  i.e. the median/mean or spread?}}
\label{fig:testHists}
\end{figure}

Results from these tests are shown in Figure \ref{fig:testHists}.  The result of the self tests for LFC exposures is shown as a black outline.  This histogram combines residuals from all exposures.  The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of the principal component amplitudes $a_{nk}$.

Overplotted in blue are the residuals from training the wavelength solution with only 90\% of LFC exposures and predicting the wavelengths for the remaining 10\% of exposures.  The two histograms match very closely, indicating that very little error is introduced when predicting denoised x values for times not used in the construction of the wavelength solution.

\section{Comparison to Classic, Parametric Model} \label{sec:comparisons}
Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.  For example, the \expres\ pipeline described in \lz{Petersburg2020} prescribed $\theta_{n}$ from equation 1 to be the amplitudes of a 2D, 9th-order polynomial in pixel $x$ and order $m$ for each exposure $n$, making the equivalent to equation 2 the following:
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\end{equation}
where $c_{nij}$ are coefficients unique to exposure $n$.  These coefficients are then smoothed by a third-order polynomial over time.  This third-order polynomial would be evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
Coefficients, $c_{nij}$, were found that minimize an objective $Q$ that is the L2-norm:
\begin{equation}
Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
\quad .
\end{equation}

Even in this simple context, where every calibration image $n$ is treated as its own unique flower, there are improvements to be made.
For one, the sums should not be from 0 to 9, but instead
\begin{equation}
\sum_{i=0}^9\sum_{j=0}^{9-i}
\quad ,
\end{equation}
because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.

For another, the objective function could be made soft to permit catastrophic outliers without destroying the fit. (\lz{we did not do this}).  We might recommend the iteratively reweighted least squares (IRLS).  This would make the fitting more robust.

For yet another, there are rescaling issues for the products $x^i\,m^j$ to protect the fitting from near-singularities or bad conditioning of the linear-algebra operators.

>>>>

\begin{abstract}
\name\ is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.  This method makes use of the many, recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.  With stabilized spectrographs, the full optical system and detectors can only vary within a tiny range of configurations.  This allows us to use all calibration data to determine the entire space of possible calibrations for a stabilized spectrograph.  A wavelength solution then need only pinpoint where in this low-dimensional calibration space the spectrograph exists.  \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.  This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.  We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.  We find that \name\ returns more accurate wavelength predictions by a factor of 10 than classic polynomial models.  Employing \name\ wavelengths reduced the RMS of RV measurements to a planet fit by \lz{some percent}.
\end{abstract}

@@

\section{Introduction} 
Extreme precision radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets (CITE THINGS).  These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.  Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).  This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrograph is well sampled).    Some hardware--software systems, such as ESPRESSO and \expres\ are now reaching these levels (CITE THINGS).

@@

The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).  We expand on the work of Milakovi\'{c}2020, which demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials.  Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).

The second idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.  Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.  That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.  This renders it inappropriate to fit each calibration exposure or calibrate each science exposure independently.  Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.  Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed for each exposure.

This structure is \emph{hierarchical}: the calibration data are used not just to determine the wavelength solution, but also to determine the possible space of wavelength solutions.  In the statistics literature this concept is often described as \emph{de-noising}.  Every calibration exposure contains information about every other exposure.  Thus every exposure can be improved (i.e. de-noised) with information from every other exposure.

The method we propose here---\name---embodies these ideas.
It is a non-parametric, hierarchical, data-driven model for the wavelength solution.  By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.  By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the observed variations in the stabilized spectrograph.  \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, etalons, or good arc lamps.  We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.  The motivating ideas behind this project are true for almost every astronomical spectrograph.

\section{Method} \label{sec:method}
\name\ is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength solution.  We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized, spectrograph hardware.

We therefore assume dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an order.  Additionally, we assume the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.

We will pose the question of wavelength calibration in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\label{eq:wsol}
\end{equation}
where $\theta_{n}$ represents the parameters describing a given exposure.

Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.  For example, the pipeline described in \lz{Petersburg2020} sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$.
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\label{eq:poly_wsol}
\end{equation}
The coefficients $c_{nij}$ are then smoothed by a third-order polynomial over time.  This third-order polynomial is evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.  Each calibration image is fit to a 2D polynomial independently.

Given a stabilized instrument, however, the spectrograph should experience only low-degree variability, meaning the calibration of any image can and should be informed by the calibration of every other image.  The calibration data themselves can be used to develop a low-dimensional basis for expressing the space of possible calibrations for a stabilized spectrograph.

If the space of all calibration possibilities is in fact $K$-dimensional (where $K$ is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} could be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\label{eq:excl_wsol}
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.

@@

\subsection{De-Noising of Calibration Frames} \label{sec:denoising}
\lz{Concerns: 1) how to best phrase the problem using variables?, 2) how in-depth is too in-depth?}

Within a given range of stability, we can use all calibration images to  1) determine the space in which the instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.  Say we have $N$ calibration exposures, $exp_n$.  The first task is to identify the full list of lines we can expect to find in a calibration image.  A line, $l_{m,\lambda}$ is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength, $\lambda$.

For each exposure, $exp_n$, every line, $l_{m,\lambda}$, has an associated fitted detector position, $x_{n,l}$, for example x-pixel in an 2D extracted order.  Fitted line centers that are missing from an exposure (e.g. because the fit failed due to noise, the line is not in its usual order, etc.) can be assigned a $NaN$ for that order instead.  Let there be $M$ lines per exposure.  We can therefore construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.

The average line position in every exposure represents the fiducial, or standard calibration of the spectrograph, $g_0(x,y,m)$.  We apply principal component analysis to the difference between this fiducial calibration and each individual line position.  The returned principal components serve as basis functions,  $g_k(x,y,m)$, expressing the possible deviations of the spectrograph from this fiducial calibration.  The magnitude of each principal component for each exposure, $a_{nk}$, represents the scalar amplitude of these deviations for each exposure.  We can now use a small number, $K$, of principal components to reconstruct a de-noised version of the line positions as formulated in Equation \ref{eq:excl_wsol}.

Missing line center measurements are replaced with de-noised estimates.  We do this iteratively until the estimates of missing line centers change by less than 0.01\%.  This process can be repeated on line centers deemed as outliers by some metric, to account for lines that may have been mis-identified or mis-fit.  The principal components from the final iteration are used to define the spectrograph's calibration space while the associated magnitudes pinpoint where in that calibration space the spectrograph is for each calibration exposure.

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $[x_{n,l}] $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
\While{change in missing or outlier line centers $>$ 0.01\%}{
	$g_0(x,y,m) = \overline{x_{n,l}]}$\;
	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
	let $a_{n,k} = U\cdot \Sigma$ and $g_k(x,y,m) = V$\;
	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ for $K=6$\;
	$[x_{n,l}] = \lambda(x,y,m,n)$ where $[x_{n,l}]$ was initially $NaN$
	}
\caption{De-Noising}
\end{algorithm}

\subsection{?Interpolating Calibration Frames?}
 \label{sec:interp_time}
The magnitude of each principal component can be interpolated with respect to housekeeping data, such as time, temperature, telescope position, etc. to determine different calibration states of the spectrograph.  The choice of what to interpolate with respect to depends on the dominant contribution to variation in the instrument.

As an example, we can interpolate the magntiudes of the principal components with respect to time to the midpoint time of non-calibration exposures to determine the calibration state of the spectrograph for that exposure.  Using interpolated magnitudes, $a'_{nk}$ and the basis vectors, $g_k(x,y,m)$, returned by the de-noising process, we can construct a set of calibration lines, $\lambda(x,y,m,n)$, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.

\subsection{Interpolating a Wavelength Solution} \label{sec:interp_wsol}
We now have the power to construct a set of calibration lines, $\{l_{m,\lambda}\}$, each with an assigned, known wavelength for any exposure.  To construct a wavelength solution, we need only interpolate the known wavelengths over the exposure specific line centers.  For instance, interpolating the known wavelengths vs. line centers onto every integer x will generate wavelengths for each pixel in an order.

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: $g_0(x,y,m)$, the fiducial calibration of the spectrograph; $a_{n,k}$, magnitudes of the principal components for each exposure; $g_k(x,y,m)$, basis vectors spanning the calibration space of the spectrograph; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
Interpolate $a_{n,k}$ with respect to time to determine $a_{n,k}'$ for time $t'$\;
Calculate pixel line positions for time $t'$ by evaluating\;
$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ for $K=6$\;
\For{each unique $m$}{
	interpolate ${l_n,m}$ onto pixels, $x'$, that we want wavelengths for\;
	}
\caption{Generating a Wavelength Solution}
\end{algorithm}

\section{Data} \label{sec:data}
We tested \name\ results using data from \expres\, the EXtreme PRecison Spectrograph.  \expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).  \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g. \lz{Wilken2012, Molaro2013, Probst2014}).

Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures every ~15-30 minutes.  ThAr exposures are taken at the beginning and end of each night.  All calibration data is taken through the science fiber, thereby traveling through the same optics and hitting the same pixels as all science observations.

The results and discussion presented here are based off of 1227 LFC exposures taken between October 14 and December 18, 2020 on 29 unique nights.  LFC lines cover 40 echelle orders, which contain 19203 lines.  Though we primarily work with LFC data, we at times discuss applications to arc lamps.  These results use 78 ThAr exposures that were taken on the same nights.  ThAr lines cover all 86 extracted orders of \expres, which includes 5295 lines.

 \name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.  We, therefore, use line positions generated by the pre-existing \expres\ pipeline, as describe in \lz{Petersburg2020}.

 A ThAr wavelength solution is generated from each ThAr exposure using the IDL code \texttt{thid.pro}, developed by Jeff Valenti.  This code identifies ThAr lines by matching lines in an exposure against a line atlas.

 Flat-relative, optimally extracted LFC data is background-corrected using a univariate spline.  Each peak in an order is then fit to a Gaussian.  The mean of the fitted Gaussian is taken to be the center of the line.  For each line, the ThAr wavelength solution is used to estimate the mode number of a line.  The precise wavelength is then calculated using
 \begin{equation}
 f_n = n \times  f_r + f_0
 \label{eq:lfc}
 \end{equation}
 where the repetition rate, $f_r$, is known from design of the LFC, and the offset frequency, $f_0$, has been determined experimentally.  the pre-existing \expres\ pipeline then generates wavelength solutions on a night-by-night basis as described in Section \ref{sec:method}.

 \lz{(Mention degeneracy with prescribed line center and wavelength solution here?  Or in discussion?)}

 We satisfy the assumption of stability needed for \name\ by using exposures from only one ``instrumental epoch" of \expres.  \expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes--changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.  Significant instrumental changes break the assumption of a stabilized instrument with only low-dimensional variations.  We therefore treat each instrument epoch independently.

\section{Tests}
We performed a series of validation tests to test 1) the de-noising step (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), 2) the interpolated wavelength solution (\textsection \ref{sec:interp_wsol}), and 3) the line density for which an interpolated wavelength solution is justified.  These tests center on the idea of predicting wavelengths for calibration lines, and comparing the predicted wavelength for each line to the assigned wavelength.

We compare the residuals of an \name -generated wavelength solution to wavelengths generated by a classic, parametric method.  Lastly, we compare RVs from pipelines employing \name\ vs. a classic wavelength calibration method.

\subsection{Test of De-Noising}\label{sec:test-denoising}
To test how well \name\ recreates calibration lines for non-calibration exposures, we employed \name\ on 90\% of all LFC exposures.  This means the basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, were constructed using only information from 90\% of all exposures.  We used the results to predict wavelengths for the lines in the remaining 10\% of calibration exposures.

\subsection{Test of Interpolation}\label{sec:test-interp}
To test how well an interpolated wavelength solution can predict wavelengths, we employed \name\ on all LFC exposures with 
of lines removed.  The resultant basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, therefore have no information about the excluded lines.  We then predict wavelengths for the excluded lines and compare these predictions to their assigned wavelengths.

\begin{figure}[h]
\centering
\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
\caption{\lz{Results of validation tests.  Histogram of residuals?  Scatter plot colored by residual value?}}
\label{fig:testHists}
\end{figure}

\subsection{Test of Line Density}\label{sec:test-density}
The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines should allow for more freedom in the resultant wavelength solution.  We test this assumption by employing \name\ on ThAr exposures.  We then use the results to predict wavelengths for the completely independent LFC exposures taken during the same range of time.

\subsection{Classic Model vs. \name } \label{sec:test-classic}
To compare the classic, polynomial-driven method of wavelength calibration and an interpolated wavelength solution, we focus on one exposure of LFC lines.  Within this exposure, we separate the even and odd lines.  We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines.

In the case of the classic model, this meant fitting a polynomial to just the odd lines and evaluating it at the position of the even lines.  In the case of an interpolated model, this mean interpolating the wavelengths of the odd lines onto the position of the even lines.  We then constructed a wavelength model using only the even lines and used it to predict wavelengths for the even lines.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{Figures/parametricTest.png}
%\includegraphics[width=\textwidth]{Figures/parametricTest.pdf}
\caption{Comparison of residuals to a parametric model (left-most) vs. a non-parametric model (right-most).  In each case, the color bar shows the difference between the predicted wavelength and the true wavelength for each line in units of $/mps$.  The center plot shows a histogram of the residuals of the polynomial fit (blue) and the non-parametric fit (green).  Medians of the residuals for each model are given in the top-right corner of the center plot in their corresponding colors.}
\label{fig:comparison}
\end{figure*}

The results to this test are shown in Figure \ref{fig:comparison}.  For each method, we plot each line with respect to its order (y-axis) and x-pixel on the detector (x-axis).  Each marker is colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.

The residuals of the classic model is shown in the left-most plot of Figure \ref{fig:comparison}; the residuals of the interpolated model is shown in the right-most plot.  The center plot is a histogram of the residuals for both methods.

The residuals of the interpolated model show much less structure, have a smaller spread, and are more symmetric about zero.  We believe the flexibility of the interpolated model was able to account for high-order detector defects, which can be seen in the structure in the residuals of the classic, smooth, polynomial model.

\subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
\lz{Debra and RP currently working through this.  Results seem good.  Will ask them at next group meeting what we should include here}

\section{Choice Choices} \label{sec:choices}
This implementation of \name incorporated determining various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.  This section will describe each choice and the associated decision-making process.
